\documentclass[11pt]{preprint}

\setlength{\topmargin}{0mm} \setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm} \setlength{\textheight}{215mm}

\usepackage{amssymb,amsmath,amscd,amsthm}

\title{Discrete Mathematics, Sect 002, 2016 Fall - Quiz 4}
\author{Name:}
\institute{Courant Institute of Mathematical Sciences, NYU}
\date{October 2, 2016}

\newtheorem*{proposition}{Proposition}



\begin{document}

\maketitle

This quiz is scheduled for 10 minutes. No outside notes or calculators are permitted. To get full credit  in all of the problems, use rigorous justification and unless otherwise indicated, make sure that your solution reads as a perfect English sentence. You should only assume the notion of integers, operations, order relations and geometrical objects as given. If you use a statement or a definition from the textbook, make sure to indicate it.
\vspace{0.2cm}

\begin{enumerate}
\item(10 points) Identify whether the following relation on the set \{1, 2, 3, 4\} is reflexive,
irreflexive, symmetric, antisymmetric, and/or transitive. Is it an equivalence relation? When the answer is no, support why.
\[
R = \{(1, 2), (3, 2), (4, 3), (2, 3), (3, 4), (2, 1)\}
\]
\begin{itemize}
\item Reflexive:
\item Irreflexive:
\item Antisymmetric:
\item Transitive:
\item Equivalence relation:
\end{itemize}
\vspace{2cm}
\item (10 points) 20 people are to be divided into two teams with ten players on each team. In how many ways can this be done? Define an equivalence relation on the set of all arrangements of the 20 people and then count how many elements are in the equivalence classes.
\end{enumerate}
\end{document}