\documentclass[11pt]{preprint}

\setlength{\topmargin}{0mm} \setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm} \setlength{\textheight}{215mm}

\usepackage{amssymb,amsmath,amscd,amsthm}

\title{Discrete Mathematics, Sect 001, 2016 Fall - Quiz 1}
\author{Name:}
\institute{Courant Institute of Mathematical Sciences, NYU}
\date{September 14, 2016}

\newtheorem*{proposition}{Proposition}



\begin{document}

\maketitle

This quiz is scheduled for 10 minutes. No outside notes or calculators are permitted. To get full credit  in all of the problems, use rigorous justification and unless otherwise indicated, make sure that your solution reads as a perfect English sentence. You should only assume the notion of integers, operations, order relations and geometrical objects as given. If you use a statement or a definition from the textbook, make sure to indicate it.
\vspace{0.2cm}

\begin{enumerate}
\item (20 points)
\begin{enumerate}
\item State the definition of divisibility.
\vspace{3cm}
\item Using the definition of divisibility above, prove or disprove the following statement:
\begin{proposition}
If $a$, $b$, $c$ are integers, and $a|b$ then $ac|bc$.
\end{proposition}
\end{enumerate}

\vspace{3cm}

\item Now switch your quiz with a person sitting next to you when ready of when the time runs out. Respectfully critique each other's work. Put both of your name on both quizzes indicating who was the original owner and hand it in together. You are allowed to make only minor adjustments, like fixing typos or phrasing but not to write down something completely new.
\end{enumerate}

\end{document}