\documentclass[11pt]{preprint}

\setlength{\topmargin}{0mm} \setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm} \setlength{\textheight}{215mm}

\usepackage{amssymb,amsmath,amscd,amsthm}

\newtheorem{proposition}{Proposition}

\title{Discrete Mathematics, 2016 Spring - Worksheet 3}
\author{Instructor: Zsolt Pajor-Gyulai, CIMS}

\date{September 14, 2016}



\begin{document}

\maketitle

In all of the above problems explain your answer in full English sentences.

\begin{enumerate}
\item Disprove the following statements:
\begin{enumerate}
\item If $a$ and $b$ are integers with $a|b$, then $a\leq b$.
\item If $a$, $b$, and $c$ are positive integers with $a|(bc)$, then $a|b$ or $a|c$.
\item If $p$ and $q$ are prime, then $p+q$ is composite.
\item Two right triangles have the same area if and only if the lengths of their hypotenuses are the same.
\end{enumerate}

\item What does it mean for an if and only if statement to be false? What properties should a counterexample for an if-and-only-if statement have?
\item Evaluate the following Boolean expressions:
\begin{enumerate}
\item $True\wedge True\wedge True\wedge True\wedge False$.
\item $(\neg True)\vee True$. 
\item $\neg(True\vee True)$. 
\item $(True\vee True)\wedge False$.
\item $True\vee(True\wedge False)$. 
\end{enumerate}
\item Prove the following Boolean identities by truth tables:
\begin{enumerate}
\item $\neg(x\wedge y)=(\neg x)\vee (\neg y)$ and $\neg(x\vee y)=(\neg x)\wedge(\neg y)$ (DeMorgan's laws).
\item $x\rightarrow y= (\neg x)\vee y$.
\item $x\leftrightarrow y = (\neg x)\leftrightarrow (\neg y)$.
\end{enumerate}
where $=$ stands for logically equivalent.
\item Find a logically equivalent Boolean expression to $x\leftrightarrow y$ only in terms of the basic Boolean operations $\wedge,\vee,\neg$.
\item How many different binary operation can there be? 
%\item (Hard problem) 
%\begin{enumerate}
%\item Show that all binary operations can be expressed in terms of the three basic Boolean operations.
%\item Show that $x\vee y$ can be expressed in terms of just $\wedge$ and $\neg$.
%\end{enumerate}
%This means that all binary Boolean operations can be reduced to just two binary operations. This fact has great implications in digital technology.

\item A person's inititals are the two-element lists consisting of the initial letters of their first and last names. For example, mines are $ZP$. 
\begin{enumerate}
\item How many possible initials are there?
\item How many initials are there where the two letters are different?
\end{enumerate}
\item A club has 10 members.
\begin{enumerate}
\item A club has 10 members who wish to elect a president and a vice-president. How many ways can these positions be filled (assuming the club is not a cult-of personality dictatorship and one person can only have one title)?
\item Now suppose the club also wants to elect a secretary and a treasurer. How many outcomes are there for the election then?
\end{enumerate}

\item In how many ways can a black rook and a white rook be placed on different squares of a chess board such that neither is attacking the other?

\vspace{0.2cm}

\noindent\underline{\textbf{Further problems to practice:}}
\vspace{0.1cm}

\item License plates in a cetrtain state consist of six characters: The first three characters are uppercase letters (A-Z), and the last three characters are digits (0-9).
\begin{enumerate}
\item How many license plates are possible?
\item How many license plates are possible if no character maybe repeated on the same plate?
\end{enumerate}

\item A telephone number (in the US and Canada) is a ten digit number whose first digit cannot be a $0$ or a $1$. How many telephone numbers are possible?

\item A US Social Security number is a nine-digit number. The first digit may be zero.
\begin{enumerate}
\item How many of these are even?
\item How many have all of their digits even?
\item How many read the same backwards? (e.g. 122979221)
\end{enumerate}

\end{enumerate}
\end{document}